TriColorings of Knots

### 2018-04-26 paper/paper.tex
Finished revising paper

\maketitle
\textbf{TODO}:
\begin{enumerate}
\item Remove \textbf{TODO}
\item Iron out Figure/image placement before final edition. Too early
  now. Wait till revision is done.
\item Passive voice/adverbs (Hemingway App)
\end{enumerate}

@@

One common pattern we see in mathematics is a concern with how
certain objects fit inside other objects. Knot theory studies this
sort of problem. Given a circle $S^1$ what different ways can we embed
it into $\bb{R}^3$ or $\bb{R}^3$  with an additional point at infinity ($S^3$)?
A typical drawing of a knot can be seen in Figure~\ref{fig:figure-eight}.

@@

Naturally, once we begin to consider knots, we are also led to consider
how we might tell them apart or rather when they should be the same
at all. Intuitively, the knot in Figure~\ref{fig:figure-eight} should
be considered the same as that in Figure~\ref{fig:figure-eight-t} but should be
different from that of trefoil seen in Figure~\ref{fig:trefoil}. In
the next section, we will see the definition of a knot, how to define their
equivalence in a sensible way, followed by a method to differentiate them
called tricolorability. Afterwards, we will look at an alternative definition
of tricolorability. Then we will introduce notions of how one may consider
coloring knots with more than three colors.

@@

Now there are two issues that we need to consider before moving forward.
The first is our notion of equivalence. Unfortunately, a homeomorphism from
one knot to another is too weak a notion of equivalence. Instead we will
use the following:

\begin{definition}
  Two knots $K,J$ are \textit{ambient isotopic} if there is a homotopy on the ambient
  space $H:S^3\times [0,1]\rightarrow S^3$ if $H(x,0)$ is the identity and the
  image $H(K,1)=J$.
\end{definition}

@@

\begin{definition}
  A \textit{diagram} $D$ for a knot $K$ is a projection of $K$ to a plane
  such that:
  \begin{itemize}
  \item At most two points are sent to the same location on the plane.
    These are called \textit{crossings}.
  \item For any two crossings there are open disjoint neighborhoods of each.
  \item All crossings must be either overcrossings or undercrossings.
  \end{itemize}
  Note that all figures in this paper are diagrams of knots.
\end{definition}

When we project to a plane there is a loss of information that occurs. This
is why we must denote all crossings as either overcrossings or undercrossings.
This allows one to recover the knot in its entirity from the diagram. Otherwise
it would be ambiguous what knot is recovered. The segments that make up the diagram
are called \textit{arcs}.

There is one other unfortunate occurrence
from moving to diagrams. It is that there are infinitely many diagrams
for a given knot. Moreover actually telling whether two diagrams denote
the same knot is extremely difficult although it is decidable and moreover
belongs to NP~\cite{Poonen}.

@@

There are two knots that we will return to over and over again in this paper
as they are the simplest examples of knots for which tricolorability
differentiates them. The simplest knot is the \textit{unknot} (see Figure~\ref{fig:unknot}), which
is unique in that it has a diagram with no crossings whatsoever. The
other is the \textit{trefoil} shown in Figure~\ref{fig:trefoil}.

@@

  For type-I we have there only a single case as a type-I move acts only on
  a single arc.
  \begin{figure}
    \includegraphics[scale=\ds]{t1-c}
    \caption{type-I}
    \label{fig:t1-c}
  \end{figure}

  For type-II, there are two cases either both strands have the same color
  or they are different.
  \begin{figure}
    \includegraphics[scale=\ds]{t2-c}
    \caption{type-II}
    \label{fig:t2-c}
  \end{figure}

  Finally, for type-III moves there are several cases based on
  the colors of the strands entering. We will demonstrate
  three such cases and the rest are similar.
  \begin{figure}
    \includegraphics[scale=\ds]{t3-c}
    \caption{type-III}
    \label{fig:t3-c}
  \end{figure}

@@

What are the benefits of looking at $3$-coloring through
the lens of Fox $3$-coloring? Unlike with coloring diagrams,
it is immediate that Fox $3$-coloring is a knot invariant
as the fundamental group is invariant under homotopy equivalence, and
so ambient isotopy as well. Another benefit
is that it is more readily apparent how we can further
extend Fox $3$-coloring than extending $3$-coloring. From
the definition of $3$-coloring, there is some ambiguity that
would need to be sorted out about precisely what rules need
to change. For example, one of our conditions was that either
all colors appear at a crossing or none. Since a crossing has
at most three arcs entering it would not be possible to have
all colors meet if we were attempting to color with more than three.

@@

\section*{Acknowledgments}
\begin{thebibliography}{99}

>>>>

\maketitle

@@

One common pattern we see in mathematics is a concern with how
certain objects fit inside other objects. Knot theory studies this
sort of problem. Given a circle $S^1$ what different ways can we embed
it into $\bb{R}^3$ or $S^3$ ($\bb{R}^3$ with a point at $\infty$)?
A typical drawing of a knot can be seen in Figure~\ref{fig:figure-eight}.

Naturally, once we begin to consider knots, we are also led to consider
how we might tell them apart or rather when they should be the same
at all. Intuitively, the knot in Figure~\ref{fig:figure-eight} should
be considered the same as that in Figure~\ref{fig:figure-eight-t} but should be
different from that of trefoil seen in Figure~\ref{fig:trefoil}. In
the next section, we will see the definition of a knot, how to define their
equivalence in a sensible way, followed by a method to differentiate them
called tricolorability. Afterwards, we will look at an alternative definition
of tricolorability that utilizes the fundamental group.
Then we will introduce notions of how one may consider
coloring knots with more than three colors.

@@

@@

Now there are two issues that we need to consider before moving forward.
The first is our notion of equivalence. Unfortunately, a homeomorphism from
one knot to another is too weak a notion of equivalence as all simple closed curves
are homeomorphic to $S^1$. Instead we will use the following:

\begin{definition}
  Two knots $K$ and $J$ are \textit{ambient isotopic} if there is a homotopy on the ambient
  space $H:S^3\times [0,1]\rightarrow S^3$ if $H(x,0)$ is the identity and the
  image $H(K,1)=J$.
\end{definition}

@@

\begin{definition}
  A \textit{diagram} $D$ for a knot $K$ is a projection of $K$ to a plane
  such that:
  \begin{itemize}
  \item At most two points are sent to the same location on the plane.
    These are called \textit{crossings}.
  \item Given any two crossings they each have an open neighborhood
    that is disjoint from the other's neighborhood.
  \item All crossings must be either overcrossings or undercrossings.
  \end{itemize}
  Note that all figures in this paper are diagrams of knots.
\end{definition}

When we project to a plane there is a loss of information that occurs. This
is why we must denote all crossings as either overcrossings or undercrossings.
This allows one to recover the knot in its entirety from the diagram. Otherwise
it would be ambiguous what knot is recovered since we would have to make a choice
of over or under for each crossing. The segments that make up the diagram
are called \textit{arcs}.

There is one other unfortunate occurrence
from moving to diagrams. It is that there are infinitely many diagrams
for a given knot. Moreover actually telling whether two diagrams denote
the same knot is extremely difficult, though decidable, and moreover
belongs to NP~\cite{Poonen}.

@@

There are two knots that we will return to repeatedly
as they are the simplest examples of knots for which tricolorability
differentiates them. The simplest knot is the \textit{unknot} (see Figure~\ref{fig:unknot}), which
is unique in that it has a diagram with no crossings whatsoever. The
other is the \textit{trefoil} shown in Figure~\ref{fig:trefoil}.

@@

  For type-I we have there only a single case as a type-I move acts only on
  a single arc as seen in Figure~\ref{fig:t1-c}.
  \begin{figure}
    \includegraphics[scale=\ds]{t1-c}
    \caption{type-I}
    \label{fig:t1-c}
  \end{figure}

  For type-II, there are two cases either both strands have the same color
  or they are different as seen in Figure~\ref{fig:t2-c}.
  \begin{figure}
    \includegraphics[scale=\ds]{t2-c}
    \caption{type-II}
    \label{fig:t2-c}
  \end{figure}

  Finally, for type-III moves there are several cases based on
  the colors of the strands entering. We will demonstrate
  three such cases, given in Figure~\ref{fig:t3-c}, and the rest are similar.
  \begin{figure}
    \includegraphics[scale=\ds]{t3-c}
    \caption{type-III}
    \label{fig:t3-c}
  \end{figure}

@@

What are the benefits of looking at $3$-coloring through
the lens of Fox $3$-coloring? Unlike with coloring diagrams,
it is immediate that Fox $3$-coloring is a knot invariant
as the fundamental group is invariant under homotopy equivalence.
So Fox $3$-coloring must be invariant under ambient isotopy as well. Another benefit
is that it is more readily apparent how we can further
extend Fox $3$-coloring than extending $3$-coloring. From
the definition of $3$-coloring, there is some ambiguity that
would need to be sorted out about precisely what rules need
to change. For example, one of our conditions was that either
all colors appear at a crossing or none. Since a crossing has
at most three arcs entering it would not be possible to have
all colors meet if we were attempting to color with more than three.

@@

\pagebreak

\begin{thebibliography}{99}
